`SimSurvey`: an `R` package to compare the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks

### 2019-12-16 analysis/paper/rebuttal_letter.Rmd

L69-73: I think it would be good to add the reference to the specific function name listed in Table2. This way, I would be more explicit and avoid possible confusion. For example, in my case, when I first read the section header “simulate abundance”, I was expecting to see the way you dealt with the spatial distribution of abundance.

*We have made the suggested change.*

L94&95&104: I just want to make sure that the bias correction has been applied to these two equations. If logN (or logZ) is normally distribution with mean mu and sigma, then the N (or Z) is lognormally distributed with mean exp(mu+sigma^2/2). Therefore, when you write “mean” in table 3 for R and Z, I hope that you modified in your actual equation for log(Z) to:
log(Z) ~ normal(log(mean)-sigma^2/2, sigma)
In other word, the mean(R) is not simply exp(meanR_log_scale). exp(meanR_log_scale) is the median of the distribution of R.

*Good catch. The answer to the question is no, bias correction was not applied. This is a matter of naming convention that we struggled with. The intent was to operate in log space for these equations and, as such, the simulations were set-up to receive log means. When developing the functions, however, we waffled on whether to use `log_mean` or `mean` as the argument name and we landed on `mean` as we figured this would be a more relatable value for users to supply. Kudos to the reviewer for noticing that we failed to consider bias-correction. To be more explicit with regards to what we "mean", we have renamed the arguments to `log_mean`.*

@@

L107: there is not enough description to understand how you converted the abundance at age to abundance at length. What were the choice of length bin groups? Every 5 cm, 10cm? Such information is also missing in the tables. Please add that. But then you have to calculate some cumulative distribution (until length l) for the length at age distribution (using the VB growth equation) and do some subtractions to be able to allocate the abundance at age to each length group.
See equation A.1.14 in here for example. https://www.nwfsc.noaa.gov/news/events/program_reviews/documents/C.2_Methot_Wetzel_SSTechnicalDescription.pdf

*Another good catch as we did not provide enough information to replicate the approach. We have added more detail on the calculations behind the sim_vonB function; turns out our approach was similar to the one described in the exemplar the reviewer provided.*

### 2019-12-16 analysis/paper/sim_survey_paper.Rmd
Improve the description of the length-age-key

$$log(l) = log(l_{\infty} - (l_{\infty} - l_0) e^{-Ka}) + \varepsilon$$

Where $l_{\infty}$ is the mean asymptotic length, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$. Abundance in discrete length groups, $l_{\mathrm{group}}^{N}$, was determined by using this formula to calculate the probability of being within a specific length group given age and then using the resultant length-age-key to convert abundance at age to abundance at length. Though some typical relationships have yet to be implemented (e.g. stock-recruitment), this formulation facilitates the simulation of age-structured populations that are dynamic enough to test an array of survey designs.

@@

| **Function call**                           | **Description**                                            | **Symbol**                          |
|:------------------------------------------- |:---------------------------------------------------------- | :---------------------------------- |
| `sim_abundance(`                            | &nbsp;                                                     | &nbsp;                              |
| `+  ages = 1:20,`                           | Ages                                                       | $a$                                 |
| `+  years = 1:20,`                          | Years                                                      | $y$                                 |
| `+  R = sim_R(log_mean = log(30000000),`    | Mean recruitment^1^                                        | $\mu_r$                             |
| `+            log_sd = 0.5),`               | Standard deviation of log-recruitment                      | $\sigma_r$                          |
| `+  Z = sim_Z(log_mean = log(0.5),`         | Mean total mortality^2^                                    | $\mu_Z$                             |
| `+            log_sd = 0.2,`                | Standard deviation of total mortality (log)                | $\sigma_Z$                          |
| `+            phi_age = 0.9,`               | Correlation across ages in error around total mortality    | $\varphi_{\delta, \mathrm{age}}$    |
| `+            phi_year = 0.5),`             | Correlation across years in error around total mortality   | $\varphi_{\delta, \mathrm{year}}$   |
| `+  growth = sim_vonB(Linf = 120,`          | Mean asymptotic length (cm)                                | $l_{\infty}$                        |
| `+                    L0 = 5,`              | Length in birth year (cm)                                  | $l_0$                               |
| `+                    K = 0.1,`             | Growth rate                                                | $K$                                 |
| `+                    log_sd = 0.1,`        | Standard deviation of the von Bertalanffy growth curve     | $\sigma_l$                          |
| `+                    length_group = 3))`   | Length group bin size for abundance at length (cm)         | $l_{\mathrm{group}}^{N}$            |

@@

The next step in the simulation is to distribute the abundance at age matrix simulated using the cohort model throughout a spatial field. Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve, applied across longitude only, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$. The grid is divided into $L_{\mathrm{div}}$ divisions (e.g. NAFO or ICES divisions) and $L_{\mathrm{strat}}$ depth-based strata. The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space via:

@@

| **Function call**                                       | **Description**                                              | **Symbol**                              |
|:------------------------------------------------------- |:------------------------------------------------------------ | :-------------------------------------- |
| `sim_distribution(`                                     | &nbsp;                                                       | &nbsp;                                  |
| `+  sim,`                                               | Simulated population from `sim_abundance`                    | &nbsp;                                  |
| `+  grid = make_grid(x_range = c(-140, 140),`           | Range of grid in the x dimension                             | &nbsp;                                  |
| `+                   y_range = c(-140, 140),`           | Range of grid in the y dimension                             | &nbsp;                                  |
| `+                   res = c(3.5, 3.5),`                | Grid resolution in x and y dimensions (km) - i.e. cell area  | $A_{\mathrm{cell}}$                     |
| `+                   shelf_depth = 200,`                | Shelf depth (m)                                              | $d_{\mathrm{shelf}}$                    |
| `+                   shelf_width = 100,`                | Shelf width (km)                                             | $w_{\mathrm{shelf}}$                    |
| `+                   depth_range = c(0, 1000),`         | Depth range from coast to slope (m)                          | $[d_{\mathrm{min}}, d_{\mathrm{max}}]$  |
| `+                   n_div = 1,`                        | Number of divisions                                          | $L_{\mathrm{div}}$                      |
| `+                   strat_splits = 2,`                 | Number of times to horizontally split strat                  | &nbsp;                                  |
| `+                   strat_breaks = seq(0, 1000, 40)),` | Series of depth breaks for defining strata                   | &nbsp;                                  |
| `+  ays_covar = sim_ays_covar(sd = 2.8,`                | Standard deviation of age-year-space distribution            | $\sigma_{\xi}$                          |
| `+                            range = 300,`             | Range of spatial correlation (km)                            | $r$                                     |
| `+                            lambda = 1,`              | Smoothness of spatial correlation                            | $\lambda$                               |
| `+                            phi_age = 0.5,`           | Correlation across ages in spatial distribution              | $\varphi_{\xi, \mathrm{age}}$           |
| `+                            phi_year = 0.9,`          | Correlation across years in spatial distribution             | $\varphi_{\xi, \mathrm{year}}$          |
| `+                            group_ages = 5:20),`      | Make space-age-year variance equal across these ages         | &nbsp;                                  |
| `+  depth_par = sim_parabola(mu = 200,`                 | Depth at which abundance is typically highest (m)            | $\mu_d$                                 |
| `+                           sigma = 70))`              | Dispersion around depth of peak abundance (m)                | $\sigma_d$                              |

@@

This work has benifited from valuable feedback from numerous colleagues, including Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury and Peter Upward. We also thank Dave Cote and Joanne Morgan for providing constructive comments on a previous version of this manuscript. This work was supported by the NSERC visiting-fellow program and Fisheries and Oceans Canada.

@@

| **Equation** | **Description** |
|:---------- |:----- |
| $L$ | Number of strata ($h = 1, 2, ....., L$) |
| $A_h$ | Area of the $h^{th}$ stratum |
| $A_{\mathrm{trawl}}$ | Area covered by a standard trawl |
| $N_h = \frac{A_h}{A_{\mathrm{trawl}}}$ | Total number of sample units in the $h^{th}$ stratum |
| $n_h$ | Total number of units sampled in the $h^{th}$ stratum ($i = 1, 2, ....., n_h$) |
| $N = \sum^{L}_{h = 1} N_h$ | Total number of sample units in the survey |
| $n = \sum^{L}_{h = 1} n_h$ | Total number of observations in the survey |
| $W_h = \frac{N_h}{N}$ | Stratum weight |
| $f_h = \frac{n_h}{N_h}$ | Sampling fraction in the $h^{th}$ stratum |
| $\overline{I}_h = \sum^{n_h}_{i = 1} \frac{I_{h, i}}{n_h}$ | Sample mean in the $h^{th}$ stratum, where $I_{h, i}$ is the $i^{th}$ observation in the $h^{th}$ stratum |
| $s^2_h = \sum^{n_h}_{i = 1} \frac{(I_{h, i} - \overline{I}_h)}{(n_h - 1)}$ | Sample variance in the $h^{th}$ stratum |
| $\overline{I} = \sum^{L}_{h = 1} W_h \overline{I}_h$ | Estimate of the population mean per unit (i.e. stratified mean catch per tow) |
| $s^2(\overline{I}) = \frac{1}{N^2} \sum^{L}_{h=1} N_h (N_h - n_h) \frac{s^2_h}{n_h}$ | Estimate of the variance of the stratified mean |
| $\widehat{I} = N \overline{I}$ | Estimate of the population total over the survey area |

>>>>

$$log(L) = log(L_{\infty} - (L_{\infty} - L_0) e^{-Ka}) + \varepsilon$$

Where $L_{\infty}$ is the mean asymptotic length, $L_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{L}^2)$. Numbers at age are distributed across discrete length groups following a lognormal distribution by calculating the probability of being in a specific length group given age. These probabilities are calculated using the standard normal cumulative density function $\Phi$ for a sequence of length groups $l$ from length 0 to 10 times the maximum predicted length $L$ at an interval of $l_{group}^{N}$:

$$ \phi_{a,l} = \Phi \left ( \frac{log(L_l)}{\sigma_{L}} \right ) - \Phi \left ( \frac{log(L_{l-1})}{\sigma_{L}} \right )$$

Though some typical relationships have yet to be implemented (e.g. stock-recruitment), this formulation facilitates the simulation of age-structured populations that are dynamic enough to test an array of survey designs.

@@

| **Function call**                           | **Description**                                            | **Symbol**                          |
|:------------------------------------------- |:---------------------------------------------------------- | :---------------------------------- |
| `sim_abundance(`                            | &nbsp;                                                     | &nbsp;                              |
| `+  ages = 1:20,`                           | Ages                                                       | $a$                                 |
| `+  years = 1:20,`                          | Years                                                      | $y$                                 |
| `+  R = sim_R(log_mean = log(30000000),`    | Mean recruitment^1^                                        | $\mu_r$                             |
| `+            log_sd = 0.5),`               | Standard deviation of log-recruitment                      | $\sigma_r$                          |
| `+  Z = sim_Z(log_mean = log(0.5),`         | Mean total mortality^2^                                    | $\mu_Z$                             |
| `+            log_sd = 0.2,`                | Standard deviation of total mortality (log)                | $\sigma_Z$                          |
| `+            phi_age = 0.9,`               | Correlation across ages in error around total mortality    | $\varphi_{\delta, \mathrm{age}}$    |
| `+            phi_year = 0.5),`             | Correlation across years in error around total mortality   | $\varphi_{\delta, \mathrm{year}}$   |
| `+  growth = sim_vonB(Linf = 120,`          | Mean asymptotic length (cm)                                | $L_{\infty}$                        |
| `+                    L0 = 5,`              | Length in birth year (cm)                                  | $L_0$                               |
| `+                    K = 0.1,`             | Growth rate                                                | $K$                                 |
| `+                    log_sd = 0.1,`        | Standard deviation of the von Bertalanffy growth curve     | $\sigma_L$                          |
| `+                    length_group = 3))`   | Length group bin size for abundance at length (cm)         | $l_{group}^{N}$                     |

@@

The next step in the simulation is to distribute the abundance at age matrix simulated using the cohort model throughout a spatial field. Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve, applied across longitude only, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$. The grid is divided into $H_{\mathrm{div}}$ divisions (e.g. NAFO or ICES divisions) and $H_{\mathrm{strat}}$ depth-based strata. The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space via:

@@

| **Function call**                                       | **Description**                                              | **Symbol**                              |
|:------------------------------------------------------- |:------------------------------------------------------------ | :-------------------------------------- |
| `sim_distribution(`                                     | &nbsp;                                                       | &nbsp;                                  |
| `+  sim,`                                               | Simulated population from `sim_abundance`                    | &nbsp;                                  |
| `+  grid = make_grid(x_range = c(-140, 140),`           | Range of grid in the x dimension                             | &nbsp;                                  |
| `+                   y_range = c(-140, 140),`           | Range of grid in the y dimension                             | &nbsp;                                  |
| `+                   res = c(3.5, 3.5),`                | Grid resolution in x and y dimensions (km) - i.e. cell area  | $A_{\mathrm{cell}}$                     |
| `+                   shelf_depth = 200,`                | Shelf depth (m)                                              | $d_{\mathrm{shelf}}$                    |
| `+                   shelf_width = 100,`                | Shelf width (km)                                             | $w_{\mathrm{shelf}}$                    |
| `+                   depth_range = c(0, 1000),`         | Depth range from coast to slope (m)                          | $[d_{\mathrm{min}}, d_{\mathrm{max}}]$  |
| `+                   n_div = 1,`                        | Number of divisions                                          | $H_{\mathrm{div}}$                      |
| `+                   strat_splits = 2,`                 | Number of times to horizontally split strat                  | &nbsp;                                  |
| `+                   strat_breaks = seq(0, 1000, 40)),` | Series of depth breaks for defining strata                   | &nbsp;                                  |
| `+  ays_covar = sim_ays_covar(sd = 2.8,`                | Standard deviation of age-year-space distribution            | $\sigma_{\xi}$                          |
| `+                            range = 300,`             | Range of spatial correlation (km)                            | $r$                                     |
| `+                            lambda = 1,`              | Smoothness of spatial correlation                            | $\lambda$                               |
| `+                            phi_age = 0.5,`           | Correlation across ages in spatial distribution              | $\varphi_{\xi, \mathrm{age}}$           |
| `+                            phi_year = 0.9,`          | Correlation across years in spatial distribution             | $\varphi_{\xi, \mathrm{year}}$          |
| `+                            group_ages = 5:20),`      | Make space-age-year variance equal across these ages         | &nbsp;                                  |
| `+  depth_par = sim_parabola(mu = 200,`                 | Depth at which abundance is typically highest (m)            | $\mu_d$                                 |
| `+                           sigma = 70))`              | Dispersion around depth of peak abundance (m)                | $\sigma_d$                              |

@@

This work has benefited from valuable feedback from numerous colleagues, including Aaron Adamack, Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Pierre Pepin, Dwayne Pittman, Don Power, Craig Purchase, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury and Peter Upward. We also thank Dave Cote and Joanne Morgan for providing constructive comments on a previous version of this manuscript. This work was supported by the NSERC visiting-fellow program and Fisheries and Oceans Canada.

@@

| **Equation** | **Description** |
|:---------- |:----- |
| $H$ | Number of strata ($h = 1, 2, ....., H$) |
| $A_h$ | Area of the $h^{th}$ stratum |
| $A_{\mathrm{trawl}}$ | Area covered by a standard trawl |
| $N_h = \frac{A_h}{A_{\mathrm{trawl}}}$ | Total number of sample units in the $h^{th}$ stratum |
| $n_h$ | Total number of units sampled in the $h^{th}$ stratum ($i = 1, 2, ....., n_h$) |
| $N = \sum^{H}_{h = 1} N_h$ | Total number of sample units in the survey |
| $n = \sum^{H}_{h = 1} n_h$ | Total number of observations in the survey |
| $W_h = \frac{N_h}{N}$ | Stratum weight |
| $f_h = \frac{n_h}{N_h}$ | Sampling fraction in the $h^{th}$ stratum |
| $\overline{I}_h = \sum^{n_h}_{i = 1} \frac{I_{h, i}}{n_h}$ | Sample mean in the $h^{th}$ stratum, where $I_{h, i}$ is the $i^{th}$ observation in the $h^{th}$ stratum |
| $s^2_h = \sum^{n_h}_{i = 1} \frac{(I_{h, i} - \overline{I}_h)}{(n_h - 1)}$ | Sample variance in the $h^{th}$ stratum |
| $\overline{I} = \sum^{H}_{h = 1} W_h \overline{I}_h$ | Estimate of the population mean per unit (i.e. stratified mean catch per tow) |
| $s^2(\overline{I}) = \frac{1}{N^2} \sum^{H}_{h=1} N_h (N_h - n_h) \frac{s^2_h}{n_h}$ | Estimate of the variance of the stratified mean |
| $\widehat{I} = N \overline{I}$ | Estimate of the population total over the survey area |
